---
title: "FFT costs"
author: "Hansjörg Neth"
date: "`r Sys.Date()`"
bibliography: [../vignettes/fft.bib]
csl: ../vignettes/apa.csl
link-citations: yes
# output: pdf_document
output:
  rmdformats::html_clean:
    code_folding: show # hide/show
    self_contained: TRUE # TRUE can cause errors with rmdformats
    toc_float: true
    toc_depth: 3
    highlight: default # textmate default kate haddock monochrome #
    lightbox: true # True by default
    fig_width: 7 # in inches
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      fig.width = 7.5, 
                      fig.height = 7.5, 
                      dpi = 100, 
                      out.width = "600px", 
                      fig.align='center', 
                      message = FALSE)

# Constants:
file_name <- "costs_FFTrees.Rmd"
```

```{r load-pkg, echo = FALSE, message = FALSE, results = 'hide'}
# Packages: 
library(FFTrees)
packageVersion("FFTrees")

library(magrittr)  # for pipe operator %>%
```


Exploring cost issues in the context of FFTs and **FFTrees**. 


# Beyond accuracy: Considering decision costs in context of FFTs 

Considering costs in the context of FFTs

## Introduction

Two distinct criteria for "good" decisions: Accuracy and cost. 
What is their relation to each other? 
Are there systematic trade-offs?

Philosophically, costs are an alternative candidate criterion to measure decision success to accuracy: 
Maximize utility, rather than consistency/rationality.
(See literature on criteria for adaptive success/survival/fitness, rather than for consistency/rationality.)


## Types of costs (in content terms)

When making decision in applied contexts, we can distinguish between different types of costs: 

- _diagnostic_ costs: cost of obtaining information.  
Example: age vs. address vs. political orientation

- _treatment_ costs: cost of an intervention.  
Example: sex vs. fertility treatment; exercise vs. operation; diet or medication.   

Possible _units_ of cost: 
money, time, ease of accessibility (e.g., based on ethical or legal norms)


## Types of costs (in model terms)

<!-- 2 types of costs: outcome vs. cues (as accuracy / frugality measures) -->

Distinguish between 2 types of costs:

1. _outcome costs_ are based on decision outcomes: What are the consequences of decisions? 
2. _cue costs_ are based on cue usage: How expensive are decisions?

Note that costs can generally be viewed as 

1. a measure of _accuracy_ (especially outcome costs)
2. a measure of _frugality_ (especially cue costs)

This ambiguity/flexibility of costs is helpful, 
as maximizing a combined measure of costs can _integrate_ both aspects. 


<!-- Costs are a function of model + data: -->

How do costs occur?

Importantly, both types of cost are _not_ a property of data or decision trees by themselves, 
but rather a function of _applying trees to data_. 
Thus, costs need to be examined in a context of both concrete data and a specific decision model.

**Example**: FFT with 1 cheap and 1 expensive cue, 2 orders:
Difference in cost depends on interaction between cues and data: 
How many cases are classified on each cue?

Create a non-trivial example (with illustrative content): 

- 67:33 classification: FFT with cheap cue first has _much lower_ cost than FFT with expensive cue first.
- 50:50 classification: FFT with cheap cue first has _lower_ cost than FFT with expensive cue first.
- 33:67 classification: FFT with cheap cue first has _higher_ cost than FFT with expensive cue first.


<!-- Alternative algorithms/competition: -->

Hypothesis regarding model comparisons: 
Costs of using FFT are usually lower than for alternative algorithms. 
However, reasons for considering costs in FFTs are also that FFTs make costs much more transparent and manipulable.


### Roles of cost (in FFTs)

Distinguish between different _roles_ of costs (in FFTs): 

- as a mere outcome measure (DV), like accuracy or frugality measures.

- as a goal or criterion for optimization/selection: 
    - for good cue thresholds (`goal.threshold`)
    - for creating trees (`goal.chase`)
    - for selecting trees (`goal`)


<!-- +++ here now +++  -->

### On optimizing costs 

1. An important insight when addressing costs is that `accuracy` can be understood as "outcome costs": 
It is computed as a function of the frequency of four classification outcomes.  
The `cost.outcomes` argument allows assigning a value to each classification outcome. 
Using values of `cost.outcomes = list(hi = 0, mi = 1, fa = 1, cr = 0)` implies that both errors have costs and are equally important. Thus, the `cost` resulting from weighting outcomes by these `cost.outcomes` is a perfect negative correlation to accuracy\ `acc`. 


2. A second and different notion of cost: _Cue cost_. 

Conceptually, costs of using cues are a measure of _frugality_ in a narrower sense than outcome costs, which are a measure of _accuracy_.

Importantly, both types of cost are a property of a tree and its decisions for some data (rather than of a tree alone). 

Note: It makes no sense to optimize `goal.threshold` for cue costs (if `cost.outcomes` are all\ 0), as every cue has a fixed cue cost (that does not change by using different thresholds). 

In practice, using `"cost"` as `goal.threshold` will only incorporate outcome costs (based on `cost.outcomes`).
However, it makes perfect sense to optimize tree construction (`goal.chase`) and select trees (`goal`) based on their cue costs or overall costs. 


# Code

## Cost types

- Is using `cost.outcomes` identical to using accuracy?

    - Do we get the _same FFTs_ when using balanced `cost.outcomes` parameters and optimizing for `cost`? 
    - What are the _correlations_ between accuracy and cost measures?

If so, then the aggregate `cost` parameter already allows integrating _accuracy_ and _cue costs_!

Test\ 1: Using `heart.train` and `heart.test` data: 

```{r cost-dec-1, eval = FALSE}
# A: Setup: HD tree by "acc": ---- 

hd_0 <- FFTrees(formula = diagnosis ~ .,           # criterion and (all) predictors
                data = heart.train,                # training data
                data.test = heart.test,            # testing data
                cost.outcomes = list(hi = 0, fa = 1, mi = 1, cr = 0), # default values
                # cost.outcomes = list(hi = 0, fa = 0, mi = 0, cr = 0),    # NEW ZERO values
                main = "HD (acc)",                 # some label
                goal = "acc",
                goal.chase = "acc",
                goal.threshold = "acc",
                sens.w = 0.6,
                max.levels = NULL,
                decision.labels = c("low risk", "high risk"), # labels for decisions
                quiet = FALSE                     # enable/suppress user feedback
)

x0 <- hd_0  # copy

# Inspect results:
summary(x0)
plot(x0, what = "cues")
x0$cues$stats$train
x0$trees$definitions
x0$trees$stats$train
x0$trees$stats$test
cor(x0$trees$stats$test$acc, x0$trees$stats$test$cost)


# B: Setting ALL goals to "cost": ----

hd_1 <- FFTrees(formula = diagnosis ~ .,           # criterion and (all) predictors
                data = heart.train,                # training data
                data.test = heart.test,            # testing data
                cost.outcomes = list(hi = 0, fa = 1, mi = 1, cr = 0), # default values
                # cost.outcomes = list(hi = 0, fa = 0, mi = 0, cr = 0),    # NEW ZERO values
                main = "HD (cost)",                # some label
                goal = "cost", 
                goal.chase = "cost",
                goal.threshold = "cost",
                decision.labels = c("low risk", "high risk"), # labels for decisions
                quiet = FALSE                      # enable/suppress user feedback
)


x1 <- hd_1  # copy

# Inspect results:
plot(x1, what = "cues")
x1$cues$stats$train
x1$trees$definitions
x1$trees$stats$test
cor(x1$trees$stats$test$acc, x1$trees$stats$test$cost)

# Note equality of cues and trees:
all.equal(x0$cues$stats$train,  x1$cues$stats$train) 
all.equal(x0$trees$definitions, x1$trees$definitions)
all.equal(x0$trees$stats$test,  x1$trees$stats$test)
```

Test\ 2: Using `titanic` data: 

```{r cost-dec-2, eval = FALSE}
# A: Setup: Titanic tree by "acc"  ---- 

set.seed(123)

t_0 <- FFTrees(formula = survived  ~ .,          # criterion and (all) predictors
               data = titanic,                   # training data
               train.p = 0.50,                   # 50:50
               cost.outcomes = list(hi = 0, mi = 1, fa = 1, cr = 0), # default values
               # cost.outcomes = list(hi = 0, mi = 0, fa = 0, cr = 0),    # NEW ZERO values
               main = "T (acc)",                 # some label
               goal = "acc", 
               goal.chase = "acc",
               goal.threshold = "acc",
               decision.labels = c("low risk", "high risk"), # labels for decisions
               quiet = FALSE                      # enable/suppress user feedback
)

x0 <- t_0  # copy

# Inspect results:
plot(x0, what = "cues")
x0$cues$stats$train
x0$trees$definitions
x0$trees$stats$test
cor(x0$trees$stats$test$acc, x0$trees$stats$test$cost)

# B: Setting ALL goals to "cost": ----

set.seed(123)

t_1 <- FFTrees(formula = survived ~ .,           # criterion and (all) predictors
               data = titanic,                   # training data
               train.p = 0.50,                   # 50:50
               cost.outcomes = list(hi = 0, fa = 1, mi = 1, cr = 0), # default values
               # cost.outcomes = list(hi = 0, fa = 0, mi = 0, cr = 0),    # NEW ZERO values
               main = "T (cost)",                # some label
               goal = "cost", 
               goal.chase = "cost",
               goal.threshold = "cost",
               decision.labels = c("low risk", "high risk"), # labels for decisions
               quiet = FALSE                      # enable/suppress user feedback
)

x1 <- t_1  # copy

# Inspect results:
plot(x1, what = "cues")
x1$cues$stats$train
x1$trees$definitions
x1$trees$stats$test
cor(x1$trees$stats$test$acc, x1$trees$stats$test$cost)

# Note equality of cues and trees:
all.equal(x0$cues$stats$train,  x1$cues$stats$train) 
all.equal(x0$trees$definitions, x1$trees$definitions)
all.equal(x0$trees$stats$test,  x1$trees$stats$test)
```

**Answer**: 
Optimizing for accuracy (without balancing or weighting sens and spec) is the same 
as optimizing for "cost" (all 3 goals) with default `cost.outcomes` values (hi = cr = 0, fa = mi = 1). 
Thus, the notions of "accuracy" and "outcome costs" are identical (provided the four outcome types are weighted the same in both measures). 

- **Caveat**: Is `goal.threshold = "cost"` actually being used in cue evaluations?


### Goal types

Note some hypotheses:

- `goal.threshold = "cost"` should affect the cue evaluations:   
If `cost.outcomes` are varied, this should be used in `fftrees_cuerank()` to alter the thresholds in `x$cues$stats$train`. 

- cue thresholds should vary with `cost.outcomes` (analogous to using `sens.w` and `wacc`), 
but _not_ vary with `cost.cues` (as every cue has a constant cost here). 

Test hypotheses: 

```{r cost-dec-3, eval = FALSE}
# A: Setup: HD tree by "acc": ---- 

h0 <- FFTrees(formula = diagnosis ~ .,           # criterion and (all) predictors
              data = heart.train,                # training data
              data.test = heart.test,            # testing data
              cost.outcomes = list(hi = 0, fa = 1, mi = 1, cr = 0),      # default values
              # cost.outcomes = list(hi = 0, fa = 0, mi = 0, cr = 0),    # NEW ZERO values
              main = "HD (acc)",                 # some label
              goal = "acc",
              goal.chase = "acc",
              goal.threshold = "acc",
              decision.labels = c("low risk", "high risk"), # labels for decisions
              quiet = FALSE                      # enable/suppress user feedback
)

# summary(h0)
h0$cues$thresholds$train$sex
h0$cues$stats


# B: Set goal.threshold to "cost" (with default `cost.outcomes`): ---- 

h1 <- FFTrees(formula = diagnosis ~ .,           # criterion and (all) predictors
              data = heart.train,                # training data
              data.test = heart.test,            # testing data
              cost.outcomes = list(hi = 0, fa = 1, mi = 1, cr = 0),      # default values !
              # cost.outcomes = list(hi = 0, fa = 0, mi = 0, cr = 0),    # NEW ZERO values
              main = "HD (cost 1)",              # some label
              goal = "acc", 
              goal.chase = "acc",
              goal.threshold = "cost",  # !
              decision.labels = c("low risk", "high risk"), # labels for decisions
              quiet = FALSE                      # enable/suppress user feedback
)

# summary(h1)
h1$cues$thresholds$train$sex
h1$cues$stats
h1$cues$thresholds$train$cp

# Note: Cue thresholds did NOT change (qed):
all.equal(h0$cues$thresholds$train$sex, h1$cues$thresholds$train$sex)
all.equal(h0$cues$stats, h1$cues$stats)


# C: Set goal.threshold to "cost" (with DIFFERENT `cost.outcomes`): ---- 

h2 <- FFTrees(formula = diagnosis ~ .,           # criterion and (all) predictors
              data = heart.train,                # training data
              data.test = heart.test,            # testing data
              # cost.outcomes = list(hi = 0, mi = 1, fa = 1, cr = 0),    # default values !
              cost.outcomes = list(hi = 0, mi = 2.5, fa = 0.5, cr = 0),    # DIFFERENT values !!
              main = "HD (cost 2)",              # some label
              goal = "acc", 
              goal.chase = "acc",
              goal.threshold = "cost", # ! Note feedback message: Using cost.outcomes, as cost.cues are constant per cue.
              decision.labels = c("low risk", "high risk"), # labels for decisions
              quiet = FALSE                      # enable/suppress user feedback
)

h2$cues$thresholds$train$sex 
h2$cues$stats  # thresholds were changed, MEAN cost is lower (but some are higher than in h1$cues$stats)

# Note: Changing cost.outcomes and optimizing goal.threshold = "cost" DID change the cue thresholds:
# - cost are different
# - order is different (as cost is used to sort rows)

all.equal(h1$cues$thresholds$train$sex, h2$cues$thresholds$train$sex) # => FALSE, qed.
all.equal(h1$cues$stats, h2$cues$stats) # => FALSE, qed.

```


- Allow setting all `cost.outcomes` to\ $0$ (to exclusively consider cue costs).

```{r cost-dec-4, eval = FALSE}
# Setup: Create FFTs by algorithm: ---- 

hd_0 <- FFTrees(formula = diagnosis ~ .,           # criterion and (all) predictors
                data = heart.train,                # training data
                data.test = heart.test,            # testing data
                # cost.outcomes = list(hi = 0, fa = 1, mi = 1, cr = 0),  # default values
                cost.outcomes = list(hi = 0, fa = 0, mi = 0, cr = 0),    # NEW ZERO values
                main = "Heart disease (0)",        # some label
                decision.labels = c("low risk", "high risk"), # labels for decisions
                quiet = FALSE                      # enable/suppress user feedback
)

x <- hd_0  # copy

# Inspect results (costs):
x$trees$stats$train


# Setting goals to "cost": ----

hd_1 <- FFTrees(formula = diagnosis ~ .,           # criterion and (all) predictors
                data = heart.train,                # training data
                data.test = heart.test,            # testing data
                # cost.outcomes = list(hi = 0, fa = 1, mi = 1, cr = 0),  # default values
                cost.outcomes = list(hi = 0, fa = 0, mi = 0, cr = 0),    # NEW ZERO values
                goal = "cost",            # works
                goal.chase = "cost",      # works
                # goal.threshold = "cost",  # FAILS, as prediction_v and criterion_v have no variance!
                main = "Heart disease (1)",        # some label
                decision.labels = c("low risk", "high risk"), # labels for decisions
                quiet = FALSE                      # enable/suppress user feedback
)

x <- hd_1  # copy

# Inspect results (costs):
x$trees$stats$train
```

**Status**: 
Setting all `cost.outcomes` to\ $0$ works, as long as not all goal parameters are set to `"cost"` as well.

However, when all `cost.outcomes` are set to\ $0$, `goal.threshold` must _not_ be `"cost"`, as cue thresholds can only be optimized for outcome costs, not cue costs (which are constant per cue).

### Adding cue costs

```{r cost-dec-5-scaling}
# Create a list with a cost value for each predictor variable:
set.seed(101)
cue_names <- names(heartdisease)[2:14]
cue_costv <- sample(1:3, size = length(cue_names), replace = TRUE)

my_cue_cost <- as.list(cue_costv)
names(my_cue_cost) <- cue_names

# Inspect cue costs:
# my_cue_cost  # list
vc_cue_cost <- unlist(my_cue_cost) # as vector
vc_cue_cost
table(vc_cue_cost)  # distribution of values?

# FFT with cue costs: ----
hd_cc1 <- FFTrees(formula = diagnosis ~ .,           # criterion and (all) predictors
                  data = heart.train,                # training data
                  data.test = heart.test,            # testing data
                  cost.outcomes = list(hi = 0, fa = 1, mi = 1, cr = 0),  # default values
                  # cost.outcomes = list(hi = 0, fa = 0, mi = 0, cr = 0),  # use ZERO values
                  cost.cues = my_cue_cost, 
                  goal = "bacc",            # works
                  goal.chase = "cost",      # set to either "bacc" OR "cast" !!!
                  goal.threshold = "bacc",  # works
                  main = "Heart disease (1)",        # some label
                  decision.labels = c("low risk", "high risk"), # labels for decisions
                  quiet = FALSE                      # enable/suppress user feedback
)

summary(hd_cc1)

```


### Combining costs

3. Define a combination: 
A weighted balance of outcome costs (accuracy) and cue costs (frugality).

Scaling issues: 
Consider scaling both components to a range from 0 (best) to 1 (worst) and then use weighted combination of both components.

```{r cost-dec-6-scaling, eval = FALSE}
# 1. Scaling/centering values: ----  

dec_cost <- c(0, 1, 1, 0)
(dc_c <- scale(dec_cost, center = T, scale = F))
dc_c/max(abs(dc_c))

cue_cost <- 1:10
cc_c <- scale(cue_cost, center = T, scale = F)
cc_c/max(abs(cc_c))


# 2. Normalizing range: Min-max scaling ---- 

d <- sample(-100:100, 50)

# normalized:
d_normalized = (d - min(d)) / (max(d) - min(d))

# # See preProcess() of caret pkg:
# d_2 <- caret::preProcess(as.data.frame(d), method = c("range"))
# d_2


# Histograms of data and normalized data:
opar <- par(no.readonly = TRUE)
par(mfrow = c(1, 2))
hist(d,            breaks = 20, xlab = "d",            col = "gold",   main = "raw data")
hist(d_normalized, breaks = 20, xlab = "Normalized d", col = "orange", main = "normalized")
par(opar)


# 3. z-transformation: ----



```


Advantage of a combined measure: Simultaneously optimizing for both accuracy and cost:

    - Compute balanced sum of "weighted outcome cost/accuracy and cue cost" (with `w.cost.outcomes` parameter from\ 0 to\ 1, default = 1 = "accuracy").
    
    - Allow using this balanced accuracy-cost as goal parameter `"cost"`.


### Visualizing trade-offs between accuracy and cost 

For a given set of FFTs (not necessarily from the same run/family): 

- Visualize training/prediction _accuracy_ (on y-axis) by _cost_ (on x-axis) in analogy to ROC curve.
- Show the marginal curve for efficient/non-dominated FFTs. 


Distinguish between 2 types of (ROC-like) efficiency curves:

1. _Accuracy_ ($y$ in range `[0, 1]`) by total _cost_ ($x$ in range `[0, Inf]`)

2. _Outcome_ costs ($y$ in range `[0, 1]`) by _cue cost_ ($x$ in range `[0, Inf]`)

<!-- Advantage:  -->

As long as the direction is correct ($0$: ideal, $-$:\ better vs. $+$:\ worse), the costs on the $x$-axis can be in any unit or scale. 

**ToDo**: 

- Identify efficient/non-dominated points from a larger set of candidate points; then 

- Use them to draw a ROC-like efficiency curve.


<!-- +++ here now +++  -->


# A grammar of FFTs

Note that `my.tree` and `fftrees_wordstofftrees()` provide a natural-language parser for defining FFTs. 
Goal: Provide simple tools for manipulating existing FFTs: 

- A grammar for editing FFTs

- A set of simple tree-trimming tools

- A **tidyverse**-friendly FFT manipulation language --- **dplyr** for FFTs ("fftplyr") 

**Workflow**: 
Start either from `FFTrees` objects or from verbal tree descriptions: 
Extract an existing FFT and then change its tree definition (in pipes). 
Use the changed definition(s) to create and evaluate new `FFTrees` objects. 

## A. Tree conversion functions

Translate between 2 FFT representations (both data frames):

1. `read_fft_df()` or "read 1 FFT" from tree definitions (was `ffts_to_fft()`): 
Interpret a line of `tree.definitions` (i.e., multiple FFTs as df) 
into a single-tree data frame (vectors of elements for each cue).  
(See `fftrees_apply()` and `fftrees_ffttowords()` for similar functionality.)

2. `write_fft_df()` or "write 1 FFT" (was: `fft_to_fft_in_1_line()`): 
Turn an FFT from single-tree df 
into a line of `tree.definitions` (as df).
(See end of `fftrees_grow_fan()` and `fftrees_wordstofftrees()` for similar functionality.)

3. `add_fft_df()` adds an FFT/its definition to a set of (existing) tree definitions. 


### Corresponding code

Create some default `FFTrees` object\ `x`:

```{r gfft-1-default, eval = TRUE}
# Setup: Create FFTs by algorithm: ---- 

hd_1 <- FFTrees(formula = diagnosis ~ .,           # criterion and (all) predictors
                data = heart.train,                # training data
                data.test = heart.test,            # testing data
                main = "Heart disease (auto)",     # some label
                decision.labels = c("low risk", "high risk"), # labels for decisions
                quiet = FALSE  # enable/suppress user feedback
)

x <- hd_1  # copy FFTrees object (with 7 FFTs)
```

Using new helper/utility functions (included in, but not yet exported by **FFTrees**):

```{r gfft-2-setup, eval = TRUE}  
# Setup: 
ffts_df_0 <- x$trees$definitions   # manually (using object structure)
ffts_df <- get_fft_definitions(x)  # using the helper function
ffts_df

# Check:
dim(ffts_df)  # 7 x 7
all.equal(ffts_df, ffts_df_0)
devtools::load_all()
```

Tree conversion functions:

Note the difference between 

- `fft`: 1 FFT (as df, with 1 row per cue)

- `ffts_df`: A (set of) tree definition(s), but also as df


```{r gfft-3-fft-conversion, eval = TRUE}
# Setup:
ffts_df <- get_fft_definitions(x)  # using the helper function
ffts_df

# Demo: 

# 1. read FFT definition: ----

fft <- read_fft_df(ffts_df = ffts_df, tree = 1)
fft  # 1 FFT (in multi-line format, as df)


# 2. write FFT definition: ----

fft_df <- write_fft_df(fft = fft, tree = 123)
fft_df  # an FFT description (as 1-line)


# 3. add FFT definition: ----

dfs_1 <- add_fft_df(fft)    # writes FFT definition and adds it to NULL
dfs_1

dfs_2 <- add_fft_df(fft_df) # adds FFT definition to NULL
dfs_2

# Intended uses:
add_fft_df(fft = ffts_df[2, ], ffts_df = dfs_2)  # 1. Adding FFT definition to a set of definitions
add_fft_df(fft = read_fft_df(ffts_df, 2), ffts_df = dfs_2)  # 2. Adding FFT (as df) to a set of definitions
```


## B. Tree editing functions 

**Goal**: Simple tools for editing and manipulating parts of existing FFTs 

Useful functions:

1. Add an FFT definition to an existing set of FFT definitions 
2. Change node orders/cue order 
3. Flip the exit direction (exit on 0/FALSE/noise vs. 1/TRUE/signal) for cue(s)
4. Change/mutate cues (by re-defining its cue, threshold, and exit direction)
5. Delete/drop/remove nodes
6. Add new nodes (assumes ability to change/mutate cues in 4. above)

- ad 1: See `add_fft_df()` above.

- ad 2: Reorder FFT cues/nodes into different orders: Sort/swap cues, change cue order, etc.  

```{r gfft-fun-2, eval = TRUE}
# Setup: 

ffts_df <- get_fft_definitions(x)  # x$trees$definitions / definitions (as df)
fft  <- read_fft_df(ffts_df, tree = 5)  # 1 FFT (as df)
fft

# Inspect FFT #5:
plot(x, tree = 5)
x$trees$definitions[5, ]  # Note: Reversed directions for exits = 0 (noise/FALSE).
x$trees$inwords[5]

# Demo: 

# Change node order:
reorder_nodes(fft)  # unchanged
reorder_nodes(fft, order = c(1, 2, 3))  # unchanged
reorder_nodes(fft, order = c(2, 1, 3))  # exit cue unchanged
reorder_nodes(fft, order = c(1, 3, 2))  # exit cue changed


# Note:
# reorder_nodes(fft, order = c(3, 2))     # ERROR: wrong length of order
```

- ad 3. Flip the exit direction/type (exit on 0/`FALSE`/noise vs. 1/`TRUE`/signal) for cue(s):
  Only flip directions of non-final nodes, as flipping the final node makes no sense? 

```{r gfft-fun-3-flip, eval = TRUE}
# Setup: 
ffts_df <- get_fft_definitions(x)  # x$trees$definitions / definitions (as df)
fft <- read_fft_df(ffts_df, tree = 2)  # 1 FFT (as df, from above)
fft

# Demo: 
flip_exits(fft, nodes = c(1))
flip_exits(fft, nodes = c(3))
flip_exits(fft, nodes = c(3, 1))
flip_exits(fft, nodes = 1:3) 

# Note: Final exits cannot be flipped:
flip_exits(fft, nodes = 4)
flip_exits(fft, nodes = 1:4)
```


## C. Macro functions 

Combinations of individual tree editing functions (with loops over sets of variations).

**Objective**: Get multiple variants of an FFT:

- Get all possible exit structures: $n$ cues correspond to $2^{n-1}$ exit structures. 

- Get all possible cue orders: $n$ cues correspond to $n!$ cue orders.


<!-- Preliminaries: -->

Require 2 utility functions on combinatorics (implemented in `util_abc.R`): 

1. `all_permutations()` yields all possible permutations of a vector\ `x`
```{r}
all_permutations(1:3) #for reordering cues (n cues haben n! Anordnungen)
all_combinations(1:3, length = 2) #for exit structure (2 ^n-1 Möglichkeiten)
```

2. `all_combinations()` yields all possible combinations of a set\ `x` of a specific\ `length`

<!-- Code: -->

**Goal**: Defining two key functions for studying FFT variants:

1. `all_node_orders()` of a given FFT:  
An FFT with $n$\ cues has $n!$ possible cue orders.

```{r gfft-fun-5-all-cue-orders, eval = TRUE}
# Setup: 
ffts_df <- get_fft_definitions(x)  # x$trees$definitions / definitions (as df)
fft  <- read_fft_df(ffts_df, tree = 1)  # 1 FFT (as df, from above)
fft

# Demo: 

# Tree 1 (3 cues): ----
dfs_1 <- all_node_orders(fft = read_fft_df(ffts_df, tree = 1))
dfs_1

dim(dfs_1)  # 3! = 6 orders

# Tree 2 (4 cues): ----
dfs_2 <- all_node_orders(fft = read_fft_df(ffts_df, tree = 2))
dfs_2

dim(dfs_2)  # 4! = 24 orders
```


2. `all_exit_structures()` of a given FFT:  
An FFT with $n$\ cues has $2^{(n-1)}$ possible exit structures.

```{r gfft-fun-6-all-exit-structures, eval = TRUE}
# Setup: 
ffts_df <- get_fft_definitions(x)  # x$trees$definitions / definitions (as df)
fft     <- read_fft_df(ffts_df, tree = 1)  # 1 FFT (as df, from above)
fft

# Demo:
dfs_3 <- all_exit_structures(fft = fft)
dfs_3

dim(dfs_3)  # 3 cues: 2^(3-1) = 4 exit structures.

dfs_4 <- all_exit_structures(fft = read_fft_df(ffts_df, tree = 2))
dfs_4

dim(dfs_4)  # 4 cues: 2^(4-1) = 8 exit structures.
```

<!-- +++ here now +++  -->

## Examples 

### Using tree editing functions

Using the auto-generated sets of FFT definitions as `tree.definitions`:

- Varying cue order:

```{r gfft-use-dfs-cue-orders, eval = FALSE}

# Evaluating ALL cue orders of dfs_1: ------  

co_1 <- FFTrees(# formula = diagnosis ~ .,           # formula
                # data = heart.train,                # training data
                # data.test = heart.test,            # testing data
                object = hd_1, 
                tree.definitions = dfs_1, 
                main = "FFT dfs_1",                 # some label
                decision.labels = c("low risk", "high risk"), # labels
                quiet = FALSE  # enable/suppress user feedback
)

summary(co_1)
plot(co_1, tree = 3, data = "train")


# Evaluating ALL cue orders of dfs_2: ------  

co_2 <- FFTrees(# formula = diagnosis ~ .,           # formula
                # data = heart.train,                # training data
                # data.test = heart.test,            # testing data
                object = hd_1, 
                tree.definitions = dfs_2, 
                main = "FFT dfs_2",                 # some label
                decision.labels = c("low risk", "high risk"), # labels
                quiet = FALSE  # enable/suppress user feedback
)

summary(co_2)
plot(co_2, tree = 24, data = "train", what = "all")
```

- Varying exit structures:

```{r gfft-use-dfs-exits, eval = FALSE}

# Evaluating ALL exit structures of dfs_3: ------  

co_3 <- FFTrees(# formula = diagnosis ~ .,           # formula
                # data = heart.train,                # training data
                # data.test = heart.test,            # testing data
                object = hd_1, 
                tree.definitions = dfs_3, 
                main = "FFT dfs_3",                 # some label
                decision.labels = c("low risk", "high risk"), # labels
                quiet = FALSE  # enable/suppress user feedback
)

summary(co_3)
plot(co_3, tree = 3, data = "train", what = "all")


# Evaluating ALL exit structures of dfs_4: ------  

co_4 <- FFTrees(# formula = diagnosis ~ .,           # formula
                # data = heart.train,                # training data
                # data.test = heart.test,            # testing data
                object = hd_1, 
                tree.definitions = dfs_4, 
                main = "FFT dfs_4",                 # some label
                decision.labels = c("low risk", "high risk"), # labels
                quiet = FALSE  # enable/suppress user feedback
)

summary(co_4)
plot(co_4, tree = 5, data = "train", what = "all")
```


More simple FFT-editing / tree-manipulation functions:

- ad 4. Change/mutate cues (by re-defining its cue, threshold, and exit direction)

- ad 5. Delete/drop/remove node(s)

- ad 6. Add node(s) (assumes ability to change/mutate cues in 4. above)


### Using tree editing functions in pipes

When working with FFTs, we can implement sequences of commands as pipes:

```{r gfft-3, eval = FALSE}
library(magrittr)  # for pipe

# 1. Start from existing tree definitions (in an FFTrees object): ------ 

# Start with an FFTrees object:
x <- hd_1  # 7 FFTs (from above)
ffts_df <- get_fft_definitions(x)  # x$trees$definitions / definitions (as df)
fft_df  <- read_fft_df_v0(ffts_df, tree = 1)  # 1 FFT (as df, from above)

plot(x, tree = 1)

# Reorder cues of tree 1:
fft_df_101 <- x %>%
  get_fft_definitions() %>%
  read_fft_df(tree = 1) %>%
  reorder_nodes(order = c(1, 3, 2)) %>%
  write_fft_df(tree = 101)
fft_df_101

# Use new tree definition:
y <- FFTrees(object = x,  # NOTE: Using arguments of objects, when none are provided.
             tree.definitions = fft_df_101, 
             main = "HD: Reorder nodes") 

plot(y, tree = 1, data = "train")


# 2. Start from a verbal description: ------

tit_0 <- FFTrees(formula = survived ~., 
                 data = titanic, main = "Titanic 0")

tit_0  # best training tree
plot(tit_0)
tit_0$trees$definitions
tit_0$trees$stats

# FFT from verbal description:  

# Check for success of Birkenhead drill:

tit_1 <- FFTrees(formula = survived ~., 
                 data = titanic, 
                 my.tree = "If sex = {female} predict True.
                            If age = {child} predict True,
                            otherwise predict False.",
                 main = "Titanic 1")

# Inspect result:
tit_1  # best training tree
plot(tit_1)
tit_1$trees$definitions
tit_1$trees$stats

# Change tree: Reverse cue order:
tit_fft_df_2 <- tit_1 %>%
  get_fft_definitions() %>%
  read_fft_df(tree = 1) %>%
  reorder_nodes(order = c(2, 1)) %>%
  write_fft_df(tree = 2)
tit_fft_df_2

# Use new tree definition:
tit_2 <- FFTrees(object = tit_1, tree.definitions = tit_fft_df_2, main = "Titanic 2")

plot(tit_2)  

# Note: Modified tree has the same 2x2 classification matrix: Same accuracy measures, 
#       but slightly worse frugality measures (as it needs to check 2 cues for adults)!

```


<!-- footer: -->

----

[File ``r file_name`` by [hn](https://neth.de/) last updated on `r Sys.Date()`]

<!-- eof. -->
